Stability of Dynamical System Ensembles

### 2014-09-10 journal.tex
add figures with constraints on char poly
remove extra comments
clarify stability condition

Consider an n-dimensional vector
$$
x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
$$
whose components are solutions to the arbitrary first order system
$$
\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
$$
where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables. Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
$$
\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
$$
where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matirx $A$ has components
$$
a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
$$
The spectral abscissa of the matrix $A$ is defined as
$$
\eta(A) = \max_i \{\Re(\lambda_i)\}
$$
where $\lambda_i$ are the eigenvalues of $A$. The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.

@@

%---------------------------------------------

% We will consider seperately the cases of $d
% < 0$ and $d > 0$.  When $d > 0$, the first condition becomes $ac/d > b$.
% Since $b > -1$, this can only be satisfied when $ac/d > -1$, i.e. when $ac > -d$.  In the subcase that $ac < d$, we then have the limits $-1 < b < ac/d$ and in the subcase $ac < d$, we have the limits $-1 < b < 1$.  When $d < 0$, the condition becomes
% $ac/d < b$, which is satisfiable when $ac > d$.  In the subcase where $ac > -d$, we have $ac/d < b < 1$ and, in the subcase where $ac < -d$, we have $-1 < b <
% 1$.

% Thus, the probability that a random stable matrix will remain stable
% upon resampling $b$ is given as follows:
% % \begin{strip}
% \begin{align*}
% P\left(\begin{pmatrix}
% a & b' \\
% d & c
% \end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
% a & b \\
% d & c
% \end{pmatrix} \textrm{stable } \right)
% & = \frac{P\left(\begin{pmatrix}
% a & b \\
% d & c
% \end{pmatrix} \textrm{stable and } \begin{pmatrix}
% a & b' \\
% d & c
% \end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
% a & b \\
% d & c
% \end{pmatrix} \textrm{stable } \right)}
% \end{align*}
% which becomes
% \begin{align*}
% &{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,
%   P(ac>b'd, a+c < 0) \over
% \int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,1} = \cr
% &{{{1 \over 32} \int_0^1 dd \int_{d > ac > -d} da\,dc \int_{-1}^{ac/d} db \,
% ({ac \over d} + 1) + {1 \over 32} \int_{0}^{1} dd \int_{ac>d} da\,dc \int_{-1}^{1} db 2 \atop +
% {1 \over 32} \int_{-1}^0 dd \int_{-d > ac > d} da\,dc \int_{ac/d}^1 db\, (1
% - {ac \over d}) + {1 \over 32} \int_{-1}^{0} dd \int_{ac>-d} da\,dc \int_{-1}^{1} db 2}
% \over
% {{{1 \over 16} \int_0^1 dd \int_{d > ac > -d} da\,dc \int_{-1}^{ac/d} db\,1
% + {1 \over 16} \int_{0}^{1} dd \int_{ac>d} da\,dc \int_{-1}^{1} db 1} \atop
% {+ {1 \over 16} \int_{-1}^0 dd \int_{-d > ac > d} da\,dc \int_{ac/d}^1 db\,1} + {1 \over 16} \int_{-1}^{0} dd \int_{ac>-d} da\,dc \int_{-1}^{1} db 1}}
% \end{align*}
% % \end{strip}

% In order to evaluate the integrals, we begin with the
% integrals over $b$, which are particularly easy because none of the
% integrands depends upon $b$.  Once we do that and make the change of
% variables $d \mapsto -d$ in the second integrals in the numerator and
% denominator, we can combine terms and simplify the expression to the
% following:
% $$
% \frac{\int_0^1 dd \int_{ac > d} da\,dc\, 4 + \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)^2}{2 \int_0^1 dd \int_{ac > d} da\,dc\, 2 + 2 \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)}
% $$

% \begin{figure}[!ht]
% \centering
% \noindent\includegraphics[width=0.5\columnwidth]{fig/2x2-resample-b.pdf}
% \caption{{\bf Resampling b.} (A) The region defined by $ac > d$ (B) The region defined by $d > ac > -d$}
% \label{fig:2x2-resample-b}
% \end{figure}

% For $ac > d$ we divide the $a$-axis into two parts indicated in Fig. \ref{fig:2x2-resample-b}A
% $$
% \int_{ac>d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{-1}^{d \over a} dc \, f(a,c) + \int_{d}^{1} da \int_{d \over a}^1 dc \, f(a,c)$$
% for the numerator $f(a,c) = 4$ and the integral is equal to 2. For $d > ac > -d$ we divide the $a$-axis into three parts indicated in Fig. \ref{fig:2x2-resample-b}B
% $$
% \int_{d > ac > -d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc \, f(a,c) + \int_{-d}^{d} da \int_{-1}^{1} dc \, f(a,c) + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc \, f(a,c)
% % \over {\int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc + \int {-d}^{d} da \int_{-1}^{1} dc + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc}
% $$
% for the numerator $f(a,c) = (\frac{ac}{d} + 1)^2$ and the integral is equal to $\frac{32}{9}$ and for the denominator $f(a,c) = \frac{ac}{d} + 1$ and the integral is equal to $3$. Therefore
% $$P\left(\begin{pmatrix}
% a & b' \\
% d & c
% \end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
% a & b \\
% d & c
% \end{pmatrix} \textrm{stable } \right) = \frac{25}{36} \approx 0.69.$$

% % Thus, our expression becomes
% % $$
% % {\int_0^1 dd \int_{-1}^{-d} da \int_{-1}^{-d/a} dc\, ({ac \over d} + 1)^2 +
% % \int_0^1 dd \int_{-d}^{d} da \int_{-1}^{1} dc\, ({ac \over d} + 1)^2 +
% % \int_0^1 dd \int_d^1 da \int_{-d/a}^{1} dc\, ({ac \over d} + 1)^2
% % \over
% % 2\int_0^1 dd \int_{-1}^{-d} da \int_{-1}^{-d/a} dc\, ({ac \over d} + 1) +
% % 2 \int_0^1 dd \int_{-d}^{d} da \int_{-1}^{1} dc\, ({ac \over d} + 1) +
% % 2 \int_0^1 dd \int_d^1 da \int_{-d/a}^{1} dc\, ({ac \over d} + 1) }
% % $$

% Now suppose we resample $a$.  Then we have to satisfy the conditions
% $a + c < 0$, $ac > bd$, and $-1 < a < 1$ for fixed values of $b,c,d$.  We will
% consider seperately the cases $c > 0$ and $c < 0$.  When $c > 0$, our
% inequalities become $a < -c$ and $a > bd/c$.  For these to be
% consistent, we must have $-c > bd/c$, or $bd < -c^2$.  In the subcase where $bd < -c$, we have $-1 < a < -c$ and, in the subcase where $-c < bd < -c^2$, we have $bd/c < a < -c$. When $c < 0$, the conditions become $a < -c$ and
% $a < bd/c$, which will have a solution as long as $bd/c > -1$, or $bd
% < -c$.  In the subcase where $bd < -c^2$, we have $-1 < a < bd/c$ and, in the subcase where $bd > -c^2$, we have $-1 < a < -c$.

% Thus, the probability that a random stable matrix will remain stable upon resampling $a$ is given as follows:
% \begin{align*}
% &\frac{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,
%   P(a'c>bd, a'+c < 0)}{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,1} = \\
% &\frac{\genfrac{}{}{0pt}{}{{1 \over 32} \int_0^1 dc \int_{bd < -c^2} db\,dd \int_{\max(bd/c,-1)}^{-c} da \,
% (-c - \max({bd \over c},-1)}{+
% {1 \over 32} \int_{-1}^0 dc \int_{bd < -c} db\,dd \int_{-1}^{\min(-c,
% bd/c)} da\, (\min(-c, bd/c) + 1)}}{{1 \over 16} \int_0^1 dc \int_{bd < -c^2} db\,dd \int_{bd/c}^{-c} da\,1
% +
% {1 \over 16} \int_{-1}^0 dc \int_{bd < -c} db\,dd \int_{-1}^{\min(-c, bd/c)} da\,1}
% \end{align*}

% \begin{figure}[!ht]
% \centering
% \noindent\includegraphics[width=0.5\columnwidth]{fig/2x2-resample-a.pdf}
% \caption{{\bf Resampling a.} (A) The region defined by $bd < -c$ (B) The region defined by $-c < bd < -c^2$ (C) The region defined by $bd < -c^2$ (D) The region defined by $bd > -c^2$}
% \label{fig:2x2-resample-a}
% \end{figure}

% For the case $c>0, {bd \over c} < a < -c, bd < -c^2$ we have two possibilities. For $bd < -c$ we have
% $$
% \int_{-1}^{-c} dd\, \int_{c/d}^1 db\, + \int_c^1 dd\, \int_{-1}^{-c/d} db,
% $$
% and for $bd > c$
% $$
% \int_{-1}^{-c} dd\, \int_{-c^2/d}^{-c/d} db\, + \int_{-c}^{-c^2} dd\, \int_{-c^2/d}^{1} db +
% \int_{c^2}^{c} dd \int_{-1}^{-c^2/d} db +
% \int_{c}^{1}dd\,\int_{-c/d}^{-c^2/d}db.
% $$
% For the case $c < 0, a < -c, a < bd/c$ we also have two possibilities. For $bd < -c^2$ we have
% $$
% \int_{-1}^{-c^2} dd\, \int_{-c^2/d}^1 db\, + \int_{c^2}^1 dd\, \int_{-1}^{-c^2/d} db,
% $$
% and for $bd > -c^2$
% $$
% \int_{-1}^{-c^2} dd\, \int_{-1}^{-c^2/d} db\, + \int_{-c^2}^{c^2} dd\, \int_{-1}^{1} db +
% \int_{c^2}^{1} dd \int_{-c^2/d}^{1} db.
% $$

%---------------------------------------------

@@

>>>>

Consider an n-dimensional vector
$$
x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
$$
whose components are solutions to the arbitrary first order system
$$
\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
$$
where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables. Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
\begin{equation}\label{eq:lineardynsys}
\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
\end{equation}
where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matirx $A$ has components
$$
a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
$$
The spectral abscissa of the matrix $A$ is defined as
$$
\eta(A) = \max_i \{\Re(\lambda_i)\}
$$
where $\lambda_i$ are the eigenvalues of $A$. The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$. This is because the general solution to equation \ref{eq:lineardynsys} is
$$
y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
$$
for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.

@@

\begin{figure}[!ht]
\centering
\noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} }
\label{fig:region2x2}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} }
\label{fig:region3x3}
\end{figure}
